Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$

### 2002-09-03 docs/paper/paper.tex
paper update
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@145 75dff457-8751-0410-ab1a-cd701654ce6b

\newcommand{\kversion}{{\tt v1\_10}}

@@

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m$ and the absolute
magnitude $M$: 
\begin{equation}
\label{kcorrecteqn}
m = M + \mathrm{DM}(z) + K(z)
\end{equation}
Sometimes a single function has been applied regardless of the galaxy
type, though more recently it has become standard to use a discrete
set of $K(z)$ functions depending on galaxy type (based either on
morphology or spectral features). As described in \cite{oke68a} and
later papers, the function is based on the projection of an assumed
galaxy SED redshifted to $z$ onto the measured instrument response as
a function of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion). For example, for the
Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
\citet{fukugita95a} have both presented results for the
$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
(called $k(z)$ by \citealt{frei94a}) for galaxies of various
morphological types, based on galaxy spectrophotometry.\footnote{Note
that the description on p. 57 of \citet{binney98a} of the meaning of
$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
exactly equivalent to our $K(z)$.}

@@

To characterize $K$-corrections, consider again Equation
(\ref{kcorrecteqn}), which we here generalize to the case of relating
an apparent magnitude observed in band $R$ to an absolute magnitude
observed in band $Q$:
\begin{equation}
m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z).
\end{equation}
The definition of an apparent AB magnitude is given by Equation
(\ref{ABdef}). An absolute AB magnitude is defined to be the apparent
AB magnitude which would be measured for a galaxy 10 pc distant, at
rest.  In the restricted case of AB magnitudes, the resulting
$K$-correction is given by:
\begin{equation}
K_{QR}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) Q(\nu)
}
\right]
\end{equation}
A more general expression holds for other magnitude systems (such as
the Vega-relative magnitude system) and for the case of
$K$-corrections when $Q$ and $R$ are in different magnitude systems.

@@

In the case of transforming $R$ to $\band{z_0}{R}$
\begin{equation}
K_{\band{z_0}{R}R}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) R(\nu/(1+z_0))
}
\right]
\end{equation}
In the special case that $z=z_0$, one is transforming the magnitude of
an observed galaxy at $z_0$ to its rest-frame magnitude in a band
shifted by $z_0$. For that case
\begin{equation}
K_{\band{z_0}{R}R}(z_0) = -2.5 \log_{10} (1+z_0)
\end{equation}
independent of $f(\nu)$ (which is natural, since both magnitudes
sample the same region of $f(\nu)$). The fact that one's uncertainties
in $f(\nu)$ do not matter makes the use of the fixed frame magnitudes
shifted to the redshift of the object in question useful, when it is
possible.

@@

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (for example, top hats). The reason is that we want the effective
bandpasses to actually have been observed for some galaxies in the
sample, so that the $K$-correction for those galaxies are formally and
actually zero. This property is desirable because, as noted above,
there are inherent uncertainties in the $K$-corrections. Nevertheless,
we note that synthesizing top hat magnitudes may be appropriate in
certain situations, as they have the virtue (similarly to the $AB$
magnitude system) that they are very easy to comprehend and synthesize
from theory.
%For example, if we wanted to convert our results into a
%form easily comprehended by the hypothetical future astronomers
%discussed in the first paragraph of the introduction, it would be best
%to calculate fluxes within wide top-hat filters so that the
%difficulties of understanding the magnitude system were minimized.

@@

Special thanks to Ivan Baldry for pointing out an error in this
manuscript and in the distributed code.  MB and DWH acknowledge NASA
NAG5-11669 for partial support.  MB was supported at the beginning of
this work by the DOE and NASA grant NAG 5-7092 at Fermilab. He is also
grateful for the hospitality of the Department of Physics and
Astronomy at the State University of New York at Stony Brook, who
kindly provided computing facilities on his frequent visits there.

>>>>

\newcommand{\kversion}{{\tt v1\_11}}

@@

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m_R$ of band $R$ and the
absolute magnitude $M$ of (in general different) band $Q$:
\begin{equation}
\label{kcorrecteqn}
m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z)
\end{equation}
The traditional definition of the $K$-correction takes $Q=R$; however,
we note that in practice many surveys do perform $K$-corrections from
one bandpass to another when comparing high redshift and low redshift
observations.

Sometimes a single function $K(z)$ has been applied regardless of the
galaxy type, though more recently it has become standard to use a
discrete set of $K(z)$ functions depending on galaxy type (based
either on morphology or spectral features). As described in
\cite{oke68a} and later papers, the function is based on the
projection of an assumed galaxy SED redshifted to $z$ onto the
measured instrument response as a function of wavelength (including
the effects of the atmosphere, the reflectivity of the mirrors, the
filter transmission, and the response of the CCD device or
photographic emulsion). For example, for the Sloan Digital Sky Survey
(SDSS) filter system, \citet{frei94a} and \citet{fukugita95a} have
both presented results for the $K$-corrections of galaxies. Each of
these papers tabulates $K(z)$ (called $k(z)$ by \citealt{frei94a}) for
galaxies of various morphological types, based on galaxy
spectrophotometry.\footnote{Note that the description on p. 57 of
\citet{binney98a} of the meaning of $k(z)$ in \citet{frei94a} is
incorrect; in fact, their $k(z)$ is exactly equivalent to our $K(z)$.}

@@

To characterize $K$-corrections, consider again Equation
(\ref{kcorrecteqn}): 
\begin{equation}
m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z).
\end{equation}
The definition of an apparent AB magnitude is given by Equation
(\ref{ABdef}). An absolute AB magnitude is defined to be the apparent
AB magnitude which would be measured for a galaxy 10 pc distant, at
rest.  In the restricted case of AB magnitudes, the resulting
$K$-correction is given by:
\begin{equation}
K_{QR}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) Q(\nu)
}
\right]
\end{equation}
A more general expression holds for other magnitude systems (such as
the Vega-relative magnitude system).

@@

In the case of transforming $R$ to $\band{z_0}{R}$
\begin{equation}
K_{\band{z_0}{R}R}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) R(\nu/(1+z_0))
}
\right]
\end{equation}
In the special case that $z=z_0$, one is transforming the magnitude of
an observed galaxy at $z_0$ to its rest-frame magnitude in a band
shifted by $z_0$. For that case
\begin{equation}
K_{\band{z_0}{R}R}(z_0) = -2.5 \log_{10} (1+z_0)
\end{equation}
independent of $f(\nu)$ (which is natural, since both magnitudes
sample the same region of $f(\nu)$). The fact that one's uncertainties
in $f(\nu)$ do not matter makes the use of the fixed frame magnitudes
shifted to the redshift of the object in question useful, when it is
possible. We note in passing that \citet{oke68a} state that the
$K$-correction ``would disappear if intensity measurements of
redshifted galaxies were made with a detector whose spectral
acceptance band was shifted by $1+z$ at all wavelengths;'' here we
have just shown this statement to be untrue for an AB system (though
one could define a magnitude system in which this were true).

@@

The reader may ask why, if we are basing our $K$-corrections on a full
model of the galaxy SED, we do not correct to bandpasses with simpler
shapes (for example, top hats). The reason is that we want the
effective bandpasses to actually have been observed for some galaxies
in the sample, so that the $K$-correction for those galaxies are
independent of the galaxy SED.  This property is desirable because, as
noted above, there are inherent uncertainties in the $K$-corrections
due to our lack of knowledge of the SEDs. Nevertheless, we note that
synthesizing top hat magnitudes may be appropriate in certain
situations, as they have the virtue (similarly to the $AB$ magnitude
system) that they are very easy to comprehend and synthesize from
theory.
%For example, if we wanted to convert our results into a
%form easily comprehended by the hypothetical future astronomers
%discussed in the first paragraph of the introduction, it would be best
%to calculate fluxes within wide top-hat filters so that the
%difficulties of understanding the magnitude system were minimized.

@@

Special thanks to Ivan Baldry for pointing out an error in this
manuscript and in the distributed code for versions {\tt v1\_10} and
previous.  MB and DWH acknowledge NASA NAG5-11669 for partial support.
MB was supported at the beginning of this work by the DOE and NASA
grant NAG 5-7092 at Fermilab. He is also grateful for the hospitality
of the Department of Physics and Astronomy at the State University of
New York at Stony Brook, who kindly provided computing facilities on
his frequent visits there.
